On the effects of partying on next-day sports performance

### 2017-07-08 writing/draft.tex
Using combined plot of party-measure by location.
Added a 'meet the spread' figure to the 'logic for using betting metrics' section. Updated tables.

\begin{figure}
  \centering
  \label{fig: next day opponent}
  \includegraphics[scale=0.45]{next_day_opponent}
  \caption{We examine the relationship between last game location and next day opponent in the NBA, sorted by distance from New York City. 
Starting from the top left, notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home. This makes sense; 1/2 of all games are played at home, and 1/2 are played away, so clearly each team is most likely to be found playing a home game for any given game. When exactly one of the last game or current game was played away,
we observe that some combinations cease to exist in our data; e.g. we never observe Charlotte Bobcats play Utah Jazz on the road and then fly directly back home for a game. Lastly, notice the patterns that emerge when both the last game and current game are played on the road. E.g. we see that when last game location was in Los Angeles, next day opponent likely to be a west coast team. There appears to be less of a correlation between last game location in New York or Brooklyn and next day opponent, perhaps because flying from east-to-west
is less burdensome as it relates to timezone changes.} 
\end{figure}

@@

@@

\begin{figure}
  \centering
  \label{fig: party by team loc}
  \includegraphics[scale=0.6,angle=270]{Party_by_Team_Location}
  \caption{We plot our continuous measure of nightlife for each city within the NBA. We present the analagous plot for MLB in the appendix.\ref{fig: party by team location mlb}}
\end{figure}

@@

% mpa1, tpa, tpo
\hspace{-125pt}
\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std. Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  
\section{Conclusions and Avenues for Future Research}
\subsection{Extending the research}
\paragraph{Within the NBA} 
We're interested to explore what other factors may affect a team's propensity to suffer hangover effects beyond simply visiting a party city; e.g. are teams more likely to go out after
a win or a loss? Do teams who aren't going to make the playoffs party more? Are there demographic
factors which influence partying, e.g. are younger, unmarried players more susceptible to treatment effect? Within a game, it would be interesting to look at 
number of shots taken, and number of shots taken by 
stars as a function of previous day visitation to party cities. We could also investigate 
effects on injury risk and flagrant foul rates.

@@

\section{Appendix}

\begin{figure}[h]
  \hspace{-55pt}
  \label{fig: party by team location mlb}
  \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
  \caption{Continuous measure of nightlife by team location for MLB.}
\end{figure}
  % What's   % interesting here is that because New York Yankees are in the Bronx, they show up % as having few musicians and sound recording studios. We might wonder if this is % a bad thing, since with NBA New York city was deemed a major party city. It % turns out that when we define a discrete indicator of party based on New York or % Los Angeles, it is not at all significant, and in fact the sign is incorrect. % When we use our continuous
%  measure of nightlife, its significant, and it helps inform our profitable % betting model.
% This is good news for our proxy, although begs the question why New York should % be treated so differently by MLB players when compared with NBA players. One % point worth noting is that we haven't yet made the important distinction that in % the MLB, the Los Angeles Angels are in Orange County, for example, which is much % less bustling than the NBA's  Los Angeles Lakers who are located in downtown Los % Angeles. 
\begin{figure}   
  \centering
  \hspace{-55pt}
  \label{fig: next day opponent mlb}
  \includegraphics[scale=0.5]{next_day_opponent_mlb} 
  \caption{For the MLB, there are some differences when compared with the     analogous
  plot for NBA. In particular, notice that a common scenario is for a team to play two away games 
  back to back at the same location. This is an artifact of baseball match-ups being played 
  in series as opposed to one-off games like in the NBA. We also notice in back-to-back
  away games that the sparse rows 
  (i.e. locations for which next-day opponent is restricted to only a strict subset of the 
  29 available teams to play) all are attributed by West coast teams. In particular, 
  Washington Nationals, San Francisco Giants, Oakland Athletics, Los Angeles Dodgers, and 
  Arizona Diamondbacks are locations from which a next-day game on the East coast are not possible.}
\end{figure}

>>>>

\begin{figure}
  \centering
  \label{fig: next day opponent}
  \includegraphics[scale=0.45]{next_day_opponent}
  \caption{We examine the relationship between last game location and next day opponent in the NBA, sorted by distance from New York City. 
Top left: notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home. This makes sense; 1/2 of all games are played at home, and 1/2 are played away, so clearly each team is most likely to be found playing a home game for any given game. When exactly one of the last game or current game was played away,
we observe that some combinations cease to exist in our data; e.g. we never observe Charlotte Bobcats play Utah Jazz on the road and then fly directly back home for a game. Lastly, notice the patterns that emerge when both the last game and current game are played on the road. E.g. we see that when last game location was in Los Angeles, next day opponent likely to be a west coast team. There appears to be less of a correlation between last game location in New York or Brooklyn and next day opponent, perhaps because flying from east-to-west
is less burdensome as it relates to timezone changes.} 
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.65]{meet_the_spread}
  \label{fig: meet the spread}
  \caption{Empirically, in NBA any given team is apt to meet the spread
half the time across a particular season; in MLB, after conditioning on bookmakers' odds of
success, successful prediction of game outcome is like a coin-flip. 
In general, systematic deviations represent
unexplained mechanisms. }
\end{figure}

@@

\begin{figure}
  \centering
  \label{fig: party by team loc}
  \includegraphics[scale=0.6,angle=270]{Party_by_Team_Location}
  \caption{We plot our continuous measure of nightlife for each city within the NBA and MLB. We add a small amount of random noise to our x-coordinate such that cities
with both an NBA and MLB team are visible.}
\end{figure}

@@

% mpa1, tpa, tpo
\hspace{-125pt}
\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std. Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}

\section{Conclusions and Avenues for Future Research}
\subsection{Extending the research}
\paragraph{Within the NBA} 
We're interested to explore what other factors may affect a team's propensity to suffer hangover effects beyond simply visiting a party city; e.g. are teams more likely to go out after
a win or a loss? Do teams who aren't going to make the playoffs party more? Are there demographic
factors which influence partying, e.g. are younger, unmarried players more susceptible to treatment effect? Within a game, it would be interesting to look at 
number of shots taken, and number of shots taken by 
stars as a function of previous day visitation to party cities. We could also investigate 
effects on injury risk and flagrant foul rates.

@@

\section{Appendix}
% \begin{figure}[h]
%   \hspace{-55pt}
%   \label{fig: party by team location mlb}
%   \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
%   \caption{Continuous measure of nightlife by team location for MLB.}
% \end{figure}
%   % What's   % interesting here is that because New York Yankees are in the % Bronx, they show up % as having few musicians and sound recording studios. We % might wonder if this is % a bad thing, since with NBA New York city was deemed a % major party city. It % turns out that when we define a discrete indicator of % party based on New York or % Los Angeles, it is not at all significant, and in % fact the sign is incorrect. % When we use our continuous
% %  measure of nightlife, its significant, and it helps inform our profitable % % betting model.
% % This is good news for our proxy, although begs the question why New York % should % be treated so differently by MLB players when compared with NBA % players. One % point worth noting is that we haven't yet made the important % distinction that in % the MLB, the Los Angeles Angels are in Orange County, for % example, which is much % less bustling than the NBA's  Los Angeles Lakers who % are located in downtown Los % Angeles. 
\begin{figure}[h]
  \centering
  \hspace{-55pt}
  \label{fig: next day opponent mlb}
  \includegraphics[scale=0.45]{next_day_opponent_mlb} 
  \caption{For the MLB, there are some differences when compared with the     analogous
  plot for NBA. In particular, notice that a common scenario is for a team to play two away games 
  back to back at the same location. This is an artifact of baseball match-ups being played 
  in series as opposed to one-off games like in the NBA. We also notice in back-to-back
  away games that the sparse rows 
  (i.e. locations for which next-day opponent is restricted to only a strict subset of the 
  29 available teams to play) all are attributed by West coast teams. In particular, 
  Washington Nationals, San Francisco Giants, Oakland Athletics, Los Angeles Dodgers, and 
  Arizona Diamondbacks are locations from which a next-day game on the East coast are not possible.}
\end{figure}
